A dream come true: deletable content in immutable storage

### 2025-09-11 archived/dream.tex
revision submitted to acm  tweb special issue

\section{Dream}

@@

In the decentralised paradigm of web 3, there is no longer a single operating entity controlling the publishing platform or the hosting infrastructure. This renders the the cost of censorship infeasible. And yet, the potential permanent exposure of personal data remains a significant concern for most individuals. Therefore, there is a need for novel solutions that can restore the sense of security that centralised gatekeepers purportedly represent.

\subsection{Deletion and revoking access}

@@

In what follows, we formulate the strongest meaningful definition of deletion applicable to decentralised storage systems and offer a construction that implements it.
Importantly, this approach is purely technical, focusing on thecapabilities and costs of primary actors, rather than relying on  procedural measures that impose obligations on intermediaries to respect the rights of primary actors.

@@

Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and the content is not accessed frequently, the chunks constituting it will be garbage-collected. However, chunks with expired postage stamps cannot be regarded as reliably deleted in order to to satisfy the requirements of sovereignty and security.

\subsection{Construction}

The goal of this section is to arrive at a formal construction of a DISC-based revocable access model. We will restrict our scheme to \emph{chunks}, the fundamental fixed-size storage units of Swarm's DISC model.  The proposed \emph{dream} construct implements a deletable content storage and access model that fulfills the requirements of specialisation, sovereignty, and security. The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model. On top of this, as is customary in Swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} expressing requirements of access control.

@@

@@

The uploader is able to construct the dream pad, which also allows them to calculate $k$ given $g$. Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$.

@@

\subsection{Analysis}

@@

\subsection{Security}

The reference does not leak any information about the step-count or the neighbourhoods involved. Neither are the participants in the protocol aware of their position in the chain or any detailsabout the other neighbourhoods, except for the immediate next one to which they are push-syncing their chunk.

@@

A powerful adversary could potentially infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it). They could also keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand. However, there is a huge cost associated with such indiscriminate archiving of all Swarm's content, and that is the only way to reliably defeat the dream construction.

>>>>

\section{Introduction}

@@

In the decentralised paradigm of web 3, there is no longer a single operating entity controlling the publishing platform or the hosting infrastructure. This renders the cost of censorship infeasible. And yet, the potential permanent exposure of personal data remains a significant concern for most individuals. Therefore, there is a need for novel solutions that can restore the sense of security that centralised gatekeepers purportedly represent.

\section{Deletion and revoking access}

@@

In what follows, we formulate the strongest meaningful definition of deletion applicable to decentralised storage systems and offer a construction that implements it.
Importantly, this approach is purely technical, focusing on the capabilities and costs of primary actors, rather than relying on  procedural measures that impose obligations on intermediaries to respect the rights of primary actors.

@@

Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and the content is not accessed frequently, the chunks constituting it will be garbage-collected. However, chunks with expired postage stamps cannot be regarded as reliably deleted in order to satisfy the requirements of sovereignty and security.

\section{Construction}

The goal of this section is to arrive at a formal construction of a DISC-based revocable access model. We will restrict our scheme to \emph{chunks}, the fundamental fixed-size storage units of Swarm's DISC model.  The proposed \emph{dream} construct implements a deletable content storage and access model that fulfils the requirements of specialisation, sovereignty, and security. The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model. On top of this, as is customary in Swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} expressing requirements of access control.

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.65\textwidth]{figs/dream.pdf}
  \caption{The dream protocoll: collective decryption of content}
\label{fig:disc}
\end{figure}

@@

The uploader is able to construct the dream pad, which also allows them to calculate $k$ given $g$. Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$, therefore the uploader can find a suitable generator $g$ in $2^d$ attempts on average.

@@

\section{Analysis}

@@

\section{Security}

The reference does not leak any information about the step-count or the neighbourhoods involved. Neither are the participants in the protocol aware of their position in the chain or any details about the other neighbourhoods, except for the immediate next one to which they are push-syncing their chunk.

@@

A powerful adversary could potentially infiltrate every neighbourhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it). They could also keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand. However, there is a huge cost associated with such indiscriminate archiving of all Swarm's content, and that is the only way to reliably defeat the dream construction.
